\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-27 coqimpl.tex
comments

To summarize, the key insight here is that, to make each recursor
handler inheritable, we need to seal an abstraction around the inductive
type (e.g. module "b_3"); but to construct a recursor (and carry out the
exhaustiveness checking), we need to break this abstraction and see the
concrete definition of the inductive type (e.g. module "b_rec_12"). Our
meta-theory need to handle these two seemingly contradicting ideas
simultaneously.

>>>>

To summarize, the key insight here is that, to make each recursor
handler inheritable, we need to seal an abstraction around the inductive
type (e.g., module "b_3"); but to construct a recursor (and carry out the
exhaustiveness checking), we need to break this abstraction and see the
concrete definition of the inductive type (e.g., module "b_rec_12"). Our
meta-theory need to handle these two seemingly contradicting ideas
simultaneously.

### 2022-06-27 syntax.tex
comments

Firstly, there are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
follows the style of~\citet{kaposi2019gluing}, but for readability we
omit universe levels attached to judgements.
After these,
we will have four new kinds of judgement to model our linkage.

Because we formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
%\YZ{What's the distinction of judgments vs. sequents?}.\EDJreply{I am not sure about the correct terminology here. Here I want to say "even underivable judgement has to be well-typed". That's the main reason I spell out the fake Agda code there. I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers. Please check}
Latex-wise, we spell out these \textit{well-formedness rule} of type
judgements and their Agda counterpart to make things more accessible. Of
course, a well-formed judgement is not necessarily derivable, just like
a well-typed QIIT type is not necessarily inhabited.

Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly---because it is not possible to have "not-well-formed" context in our setting.

Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.

@@

\begin{figure}[!htb]
  \begin{align*}
    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}
  \end{align*}
\caption{Judgements and their Agda Encoding}
\end{figure}

\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Tm]
      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
      {\goodTerm{\Gamma}{\_}{T}}
      $$
      $$
      \Rule[name=Tms]
      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Con   : Set 
      data Ty    : Con → Set   
      data Tms   : Con → Con → Set 
      data Tm    : (Γ : Con) → Ty Γ  → Set 
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

\caption{Well-formedness of Four New Judgements}
\end{figure}

@@

Now we show selected typing rules: we will have dependent function types
and dependent pair types. Dependent pair type is used to model Coq
module, which is useful when we need to model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type.

@@

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU_j[\gamma] \equiv \cU_j }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[name=Univ-1]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU_j}
}\quad
\Rule[name=Univ-2]
{\goodTerm{\Gamma}{T}{\cU_j}}
{\goodType{\Gamma}{El \ T}{j}}
\quad
\Rule
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

@@

We have two kinds of signatures, one for inductive type, the other for linkages.

@@

Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type. To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}. Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig-Proj} to extract and corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece of the syntax. Here $\bW \sigma$ is a large singleton type~\cite{stone2000}, and it only has one ``element'', the inductive type itself, inside. We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e. if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type. They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e. k |- q : (W T) is not allowed, we only have k |- q : El (W T). This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing. BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general. I just didn't write out the two equations c (El T) = T, El (c T) = T}. We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is not allowed to locate at type position (i.e. at the right-most position of a term judgement) but $El \ T$ is.

The formulation of our recursor is closely related to the linkage, so we will postpone their description.

@@

$$
\Rule[name=Ept Sig]
{}
{\goodSig{\Gamma}{\nu\cdot}{0}}
\quad
\Rule[name=Sig Add]
{\goodSig{\Gamma}{\sigma}{n} 
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}
 \quad \goodType{\Gamma, A}{T}{i}}
{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
$$

@@

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields. We omit the naming/label information for fields in the meta-theory. We can use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).

The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.  They are both constructed from the empty stuff, and respect substitution lemma for sure. We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony~\cite{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules). We omit them and the substitution laws for either field addition and projection. For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This omission can be supported by Agda using Implicit Arguments}

@@

$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\sigma}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
\quad n < N
}
{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
$$
$$
\Rule[name=Rec-Constr]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodType{\Gamma}{T}{\bW \sigma}
\quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
$$

In our meta-theory, we only formulate non-dependent eliminator, i.e. the
\textit{recursor}. Each recursor is constructed by a linkage containing
functions/handlers/pattern-matching cases dealing with each
constructor. Thus code reuse for recursor is delegated to linkage
inheritance----to reuse one particular pattern-matching case,
we just inherit the function/handler. This greatly simplifies the
metatheory development.

Intuitively \ruleref{Hdler} is indicating the type of recursor handler for one specific constructor, and \ruleref{Hdlers} is the linkage type of the bundle of all recursor handlers of one given inductive type. We omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its (apparent) internal definition. "prjR" is also the handler selection from the handler linkage, and is helpful for the formulation of $\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is the used to construct a concrete recursor given a specific linkage of handlers. All these rules satisfy substitution laws and omitted here.

@@

\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a short hand of the other)}

\begin{minted}[]{agda}
Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]

@@

As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
a short hand of term judgement $\goodTerm{\Gamma , \cC
\sigma}{\_}{A[\pi_1]}$, a special form of \textit{local morphism}~\cite{abbott2003category}. The substitution law should be understood as a
lemma, and it is easily derivable using $\gamma^\uparrow$.

The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance---due to the presence of inductive type in the context, not everything is inheritable. For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type. In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$. This makes all the fields after this inductive type non-inheritable---which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.

Sealing is used to exploit this fact---with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$. Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.

>>>>

Firstly, there are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
follows the style of~\citet{kaposi2019gluing}, but for readability we
omit universe levels attached to judgements.
After these,
we will have four new kinds of judgements to model our linkages.

Because we formalize our language as a deep embedding in a type theory with QIIT types, every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms (otherwise how can they be encoded in Agda?)
%\YZ{What's the distinction of judgments vs. sequents?}.\EDJreply{I am not sure about the correct terminology here. Here I want to say "even underivable judgement has to be well-typed". That's the main reason I spell out the fake Agda code there. I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers. Please check}
Latex-wise, we spell out these \textit{well-formedness rules} of type
judgements and their Agda counterparts to make things more accessible. Of
course, a well-formed judgement is not necessarily derivable, just like
a well-typed QIIT type is not necessarily inhabited.

Due to our intrinsic setting, \textbf{we can and will omit a lot of
presumptions of the typing rules}. For example, we don't spell out that
the rule "Type Universe" requires a well-typed context, even though it
does implicitly, because it is not possible to have ill-formed
contexts in our intrinsic setting.

Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the
most special one because it is actually just a shorthand of term
judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.\YZ{You haven't introduced the notation A[pi_1] yet at this point. And what is A[pi_1]?}

@@

\begin{figure}[!htb]
  \begin{align*}
    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over subsitutions and s to range over W-types, which add to the confusion.}
  \end{align*}
\caption{Judgement forms, as Agda types and as presented in this section.}
\end{figure}

\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Tm]
      {\goodCtx{\Gamma}{i} \quad \goodType{\Gamma}{T}{j}}
      {\goodTerm{\Gamma}{\_}{T}}
      $$
      $$
      \Rule[name=Tms]
      {\goodCtx{\Gamma}{i} \quad \goodCtx{\Delta}{i}}
      {\goodSub{\Gamma}{\_}{\Delta} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments? Besides, are these rules used anywhere in the following? Seems a bit unusual to have well-formedness rules for judgments per se.}
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Con   : Set 
      data Ty    : Con → Set   
      data Tms   : Con → Con → Set 
      data Tm    : (Γ : Con) → Ty Γ  → Set 
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

\caption{Well-formedness rules of four new judgements}
\end{figure}

@@

Now we show selected typing rules: we will have dependent function types
and dependent pair types. Dependent pair type is used to model Coq
module, which is useful when we need to model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type.
The substitution law for these two
require \textit{de bruijn lifting}: for
$\goodTerm{\Gamma}{\sigma}{\Delta}$, we define its de Bruijn lifting
$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow}{\Delta, A}$ given
$\goodType{\Delta}{A}{}$.\YZ{What is A[sigma] and what is de Bruijn lifting? Seems like if the reader is not familiar with the substitution calculus, then they will get lost starting here. Broadly speaking, a gentle intro to the intrinsic setting would help the reader a great deal.}

@@

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU_j[\gamma] \equiv \cU_j }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[name=Univ-1]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU_j}\YZ{what is c? Why is this rule called Univ-1?Didn't you write that you omit universe levels?}
}\quad
\Rule[name=Univ-2]
{\goodTerm{\Gamma}{T}{\cU_j}}
{\goodType{\Gamma}{El \ T}{j}\YZ{'El T' isn't explained only until it makes a second appearance in Ind-Term.}}
\quad
\Rule
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

@@

We have two kinds of signatures, one for inductive types, the other for linkages.

@@

Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A
\vdash B(x)$ decides a W-type. To simulate extensible inductive types, we
have to equip our inductive types with multiple constructors, and for
simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ :
(x : Aᵢ) → (Bᵢ x → W) → W}.
Thus, we use a list of pairs ($A_i, B_i$) as the signature of one
inductive type, and given an inductive type signature, we use
\ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece
of the syntax. Here $\bW \sigma$ is a large singleton
type~\cite{stone2000}, and it only has one ``element'', the inductive
type itself, inside. We formulate in this cumbersome way because we want
to have inductive type to be a field element and exposing the
constructors at the type level for later \textit{pattern-matching
exhaustiveness checking} (i.e. if an inductive type as field element has type
$\cU$, then no information about the inductive type is exposed).
\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type. They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e. k |- q : (W T) is not allowed, we only have k |- q : El (W T). This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing. BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general. I just didn't write out the two equations c (El T) = T, El (c T) = T}\YZreply{Seems harmless to list these equations and move the reference to 'type universe a la Tarski' earlier.}
We follow the style of \textit{type universe à la Tarski} and
distinguish types from their \textit{codes or names}, and thus we use
$El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is
not allowed to locate at type position (i.e. at the right-most position
of a term judgement) but $El \ T$ is.

The formulation of our recursor is closely related to that of linkages,
so we will postpone their description.

@@

$$
\Rule[name=Ept Sig]
{}
{\goodSig{\Gamma}{\nu\cdot}{0}}
\quad
\Rule[name=Sig Add]
{\goodSig{\Gamma}{\sigma}{n} 
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}\YZ{How should I understand 'f' and 'A'?}
 \quad \goodType{\Gamma, A}{T}{i}}
{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
$$

@@

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
for linkages in the context $\Gamma$ with $n$ fields. We omit the
naming/label information for fields in the meta\-theory. We can use $\cL$
to retrieve the corresponding types of linkage of a given signature, and
$\cC$ to get the compiled linkage type (an existential type).\YZ{Why is Csigma needed? How is it different from Lsigma?}

The signature (for the linkage) and the linkage rules are quite similar to
the signature for inductive types.  They are both constructed from the
empty stuff, and respect substitution lemma for sure. We have
\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
due to harmony~\cite{pfenning2009lecture}, so we have local soundness
and local completeness ($\beta,\eta$ rules). We omit them and the
substitution laws for either field addition and projection. For
$\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This
omission can be supported by Agda using Implicit Arguments}

@@

$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}}
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\sigma}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
\quad n < N
}
{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
$$
$$
\Rule[name=Rec-Constr]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodType{\Gamma}{T}{\bW \sigma}
\quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
$$

In our meta-theory, we only formulate non-dependent eliminator, i.e., the
\textit{recursor}.\YZ{Do dependent eliminators pose technical difficulties, or do you avoid them solely because they require more verbiage?}
Each recursor is constructed by a linkage containing
functions/handlers/pattern-matching cases dealing with each
constructor. Thus code reuse for recursor is delegated to linkage
inheritance---to reuse one particular pattern-matching case,
we just inherit the function/handler. This greatly simplifies the
metatheory development.

Intuitively \ruleref{Hdler} is indicating the type of recursor handler
for one specific constructor, and \ruleref{Hdlers} is the linkage type
of the bundle of all recursor handlers of one given inductive type. We
omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its
(apparent) internal definition. "prjR" is also the handler selection
from the handler linkage, and is helpful for the formulation of
$\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
the used to construct a concrete recursor given a specific linkage of
handlers. All these rules satisfy substitution laws and omitted here.
\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!). I think it's better to list all the rules and decide later what to throw to the appendix.}

@@

\caption{Seal Judgement, and its Agda Representation \\ (or What do we mean when we say one judgement is a shorthand of the other)}

\begin{minted}[escapeinside=@@]{agda}
Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]@\YZ{tau?}@

@@

As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
a shorthand of term judgement $\goodTerm{\Gamma , \cC
\sigma}{\_}{A[\pi_1]}$,\YZ{Not quite. Previously you used cL I think.}
a special form of \textit{local morphism}~\cite{abbott2003category}.
The substitution law should be understood as a lemma, and it is easily
derivable using $\gamma^\uparrow$.

The motivation for sealing judgement is acting like syntactic sugar
helpful for inheritance---due to the presence of inductive type in the
context, not everything is inheritable. For example, $ X_1 : \bW
\sigma_1 \vdash t : T$ in the parent family will generally not be able
to be inherited to the children once we override the inductive type. In
other words, there is no way to ``transplant'' $t$ into the context $X_2
: \bW \sigma_2$.\YZ{Is sigma_2 the overriding of sigma_1? The reader has to guess.}
This makes all the fields after this inductive type
non-inheritable---which shouldn't be the case because a lot of fields
might not be directly relying on this concrete $\bW \sigma_1$.

Sealing\YZ{It occurs to me that you use the word "sealing" to mean different (though tangentially related) things. Previously you used it as in "sealed families".} is used to exploit this fact---with sealing, given $X_1 : \bW
\sigma_1$, we can add well-typed terms $X_1 : \cU \vdash t : T$, into
the parent family. Later when we want to inherit $t$,
because $t$ only consider $X_1$ as an arbitrary type, inheritance is
possible.
